Microclimatic spatial variation buffers local warming in alpine vegetation

### 2022-01-14 doc/manuscript.Rmd
changes to ms

---
title: "Microclimatic spatial variation buffers local warming in alpine vegetation"
output: html_document
---

@@

B Jiménez-Alfaro, E Fernández-Pascual, A Bueno, C Marcenó….¿?

@@

The study was conducted in the calcareous massif of Picos de Europa National Park, in northern Spain (Figure 1A). The study area is a biodiversity hotspot for cold-adapted plants in the Iberian Peninsula and a biogeographical hub for Alpine and Mediterranean lineages in Western Europe (XXX). The main calcareous massif occupies c. 50 km2 and support high diversity of ecosystems, with alpine vegetation mostly occurring between 1900 and 2400 m a.s.l., with a local species pool of XX species (Jiménez-Alfaro et al. XXX). In 2008, we established a long-term ecological research program for monitoring soil climate and vegetation change. We selected four study sites along a north-south gradient, reflecting the main macroclimatic variation from oceanic to Mediterranean influence. In 2016, we additionally installed a Water potential logger (XXX) to characterize water stress in the four sites and to ….  The sites were selected to… general design and description of sites… (Borja)

@@

In each site, we buried a temperature logger (XXXX) at 5 cm depth  in a relatively flat and homogeneous vegetation patch. We surveyed the plant community in two replicated plots of 1 m2 separated 1 m from the logger, identifying species composition and estimating relative cover in %. In each plot, we installed a grid template of 100 microplots (10 x 10 cm each) to sample species frequency according to the standard methodology of GLORIA (XXX). The loggers were replaced by new ones, when needed, to obtain a continuous temperature record from 2008 to 2018. In 2018, we resampled the same plots in the same way to detect potential changes in species presence and frequency. The vegetation data from these surveys, together with the soil temperature collected in the four study sites during 10 years, represent the “temporal survey”. In 2019, we visited the same areas to study the spatial variation of vegetation and microclimate around the previously sampled areas. Using the long-term temperature logger as the central point, we additionally placed 20 ibuttons (XXX) in 20 plots of 1 m2 separated 10 m from each other in the four cardinal directions (Figure 1C).. etc… (Borja)

@@

@@

Along the 10 years of soil temperature monitoring, we found a general trend of increasing temperature, especially in the southwest and warmest site (Hoyo sin Tierra) (Figure 3). Two of the sites (Los Cazadores and Los Boches) showed a consistent pattern of continuous snow cover during winter (reflected by temperature records around 0º C). In contrast, the two other sites (Hoy Sin Tierri and Hoyo Sin Tierra) showed frost temperatures during most winters (although they may not be constant and intermixed with short snow periods). Maximum temperatures were… Soil temperature corresponding to the spatial surveys also showed high variation within sites (Figure 3). In the four sites we found plots with contrasting patterns of snow cover and frost during winter, reflected by…  When looking at the whole microclimatic variation among plots (Figure 4), we found that the four sites were represented by a wide range of temperatures, with the exception of one site (Los Cazadores). The two main axes of variation were mainly related to FDD and GDD, the latter also correlated with…. (Eduardo)

@@

In the temporal survey (2 plots per site, n =16), we recorded XX species in 2008 and XX in 2018. (species loss, gains)… Changes in species frequency were/were not correlated with XX and XX…. (Table 1) (Eduardo). In the spatial survey (20 plots per site, n = 80; including four plots placed in rocks with no plants) we recorded XX species. Changes in species composition were associated to GDD (NMDS axis 1) to represent a main gradient from the warmest to the coldest sites (Figure 4), and to FDD (NMDS2) in a gradient of snow versus frost winters. In the compositional space, the sites were poorly differentiated across the four sites (PERMANOVA, adonis) mainly by the effect of one site (Los Boches) with most plots places in the cold and frost margins. Etc…. (Borja).

@@

…. (Eduardo) (quizás repensar esta parte, pero es importante mantenerla porque es la clave).

>>>>

---
title: "Microclimatic spatial variation buffers local warming in alpine vegetation"
output: github_document
bibliography: refs.bib
csl: annals-of-botany.csl
link-citations: yes
always_allow_html: true
---

@@

B Jiménez-Alfaro, E Fernández-Pascual, A Bueno, C Marcenó....¿?

@@

The study was conducted in the calcareous massif of Picos de Europa National Park, in northern Spain (Figure 1A). The study area is a biodiversity hotspot for cold-adapted plants in the Iberian Peninsula and a biogeographical hub for Alpine and Mediterranean lineages in Western Europe (XXX). The main calcareous massif occupies c. 50 km2 and support high diversity of ecosystems, with alpine vegetation mostly occurring between 1900 and 2400 m a.s.l., with a local species pool of XX species (Jiménez-Alfaro et al. XXX). In 2008, we established a long-term ecological research program for monitoring soil climate and vegetation change. We selected four study sites along a north-south gradient, reflecting the main macroclimatic variation from oceanic to Mediterranean influence. In 2016, we additionally installed a Water potential logger (XXX) to characterize water stress in the four sites and to .... The sites were selected to... general design and description of sites... (Borja)

@@

In each site, we buried a temperature logger (XXXX) at 5 cm depth in a relatively flat and homogeneous vegetation patch. We surveyed the plant community in two replicated plots of 1 m2 separated 1 m from the logger, identifying species composition and estimating relative cover in %. In each plot, we installed a grid template of 100 microplots (10 x 10 cm each) to sample species frequency according to the standard methodology of GLORIA (XXX). The loggers were replaced by new ones, when needed, to obtain a continuous temperature record from 2008 to 2018. In 2018, we resampled the same plots in the same way to detect potential changes in species presence and frequency. The vegetation data from these surveys, together with the soil temperature collected in the four study sites during 10 years, represent the "temporal survey". In 2019, we visited the same areas to study the spatial variation of vegetation and microclimate around the previously sampled areas. Using the long-term temperature logger as the central point, we additionally placed 20 ibuttons (XXX) in 20 plots of 1 m2 separated 10 m from each other in the four cardinal directions (Figure 1C).. etc... (Borja)

@@

[@RN2315]

@@

Along the 10 years of soil temperature monitoring, we found a general trend of increasing temperature, especially in the southwest and warmest site (Hoyo sin Tierra) (Figure 3). Two of the sites (Los Cazadores and Los Boches) showed a consistent pattern of continuous snow cover during winter (reflected by temperature records around 0º C). In contrast, the two other sites (Hoy Sin Tierri and Hoyo Sin Tierra) showed frost temperatures during most winters (although they may not be constant and intermixed with short snow periods). Maximum temperatures were... Soil temperature corresponding to the spatial surveys also showed high variation within sites (Figure 3). In the four sites we found plots with contrasting patterns of snow cover and frost during winter, reflected by... When looking at the whole microclimatic variation among plots (Figure 4), we found that the four sites were represented by a wide range of temperatures, with the exception of one site (Los Cazadores). The two main axes of variation were mainly related to FDD and GDD, the latter also correlated with.... (Eduardo)

@@

In the temporal survey (2 plots per site, n =16), we recorded XX species in 2008 and XX in 2018. (species loss, gains)... Changes in species frequency were/were not correlated with XX and XX.... (Table 1) (Eduardo). In the spatial survey (20 plots per site, n = 80; including four plots placed in rocks with no plants) we recorded XX species. Changes in species composition were associated to GDD (NMDS axis 1) to represent a main gradient from the warmest to the coldest sites (Figure 4), and to FDD (NMDS2) in a gradient of snow versus frost winters. In the compositional space, the sites were poorly differentiated across the four sites (PERMANOVA, adonis) mainly by the effect of one site (Los Boches) with most plots places in the cold and frost margins. Etc.... (Borja).

@@

.... (Eduardo) (quizás repensar esta parte, pero es importante mantenerla porque es la clave).
